Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-17 chi2012.tex
all but survey

\subsection{Condition B: Observations and Issues (Websites + Excel)}
In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.  For B trials, we measured the fraction of the trial time spent in Excel versus looking at the sites in a 5-value range (0,25,50,75,100\%).  Six individuals used Excel entirely without consulting the web sites at all, while one avoided Excel altogether; most fell somewhere in between, with a median use of 75\% Excel.   When asked to explain their choice post-hoc, participants who used Excel explained that it saved them the effort of having to individually seach and consolidate information from the websites (10 participants).  Participants who looked at websites chose to do so when they needed to know more than what was provided in the spreadsheets (8 participants), when they mistrusted the information in them (1 participant), or if they were not familiar with Excel (1 participant).

Of the Excel use, most participants made minor formatting and column sizing changes to facilitate reading and interpreting the spreadsheets, while four made extensive edits to the sheets to help them ``think through'' the task in various ways.  These uses of Excel for external cognition varied from merely copying over candidate choices from one sheet to the next, sorting values by multiple columns (which participants had considerable trouble with), and colour-coding cells to allow easy counting of  ``pros'' and ``cons''; ``I can't deal with so many textual values; I need colours to help me see'' (p6).

Participants who opted to use the spreadsheet to support their decisions switched between worksheets to view data from different websites.  They used different approaches to using the spreadsheets, some participants were keen to collate all data onto a single sheet, so that they could compare the values all at once. However, this approach was error prone, for example participants noticing too late that the ordering of rows was different on each sheet, so that their pasted values were incorrectly mapped. Others used pen and paper to collate the values, which was often faster and resulted in less observed frustration. While the majority of users did not
attempt anything complex with Excel (in fact only one user used charts), there was some advanced use of Excel observed.
One user explained that they did not like to look at numbers, and so coloured the cells of the types of restaurant, and the
friends' preferences, and matched the colours to make their decision. Another user calculated their own weighted metric
of which university to choose by averaging different metrics supplied by the data sources.

>>>>

\subsection{Condition B: Observations and Issues (Websites + Excel)}
In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.  For B trials, we measured the fraction of the trial time spent in Excel versus looking at the sites in a 5-value range (0,25,50,75,100\%).  Six individuals used Excel entirely without consulting the web sites at all, while one avoided Excel altogether; most fell somewhere in between, with a median use of 75\% Excel.   When asked to explain their choice post-hoc, participants who used Excel explained that it saved them the effort of having to individually search and consolidate information from the websites (10 participants).  Participants who looked at websites chose to do so when they needed to know more than what was provided in the spreadsheets (8 participants), when they mistrusted the information in them (1 participant), or if they were not familiar with Excel (1 participant).

Of the Excel use, most participants made minor formatting and column sizing changes to facilitate reading and interpreting the spreadsheets, while four made extensive edits to the sheets to help them ``think through'' the task in various ways.  These uses of Excel for external cognition varied from merely copying over candidate choices from one sheet to the next to keep track of their previous selections, to sorting values by multiple columns (which participants had considerable trouble with due to the need to sort columns together), and colour-coding cells to allow easy counting of  ``pros'' and ``cons''; ``I can't deal with so many textual values; I need colours to help me see'' (p6). Another user calculated their own weighted metric of which university to choose by averaging different metrics supplied by the data sources and filling in their own chosen weights.

Participants ran into a number of difficulties with using the Excel interface. Participants struggled with switching between different worksheets, and having to find the corresponding values in each.  To cope with this fragmentation, some participants attempted to collate all data onto a single sheet, so that they could compare the values all at once as just described; however, this approach was often error prone.  For example, two participants did not notice that the rows in two sheets were in different orders (because of a difference in the renderings of Unviersity names), which caused their pasted values to be incorrectly mapped.  Four participants struggled with sorting columns, and only 1 attempted charting (p7); Excel's charting features caused her considerable difficulty and she eventually gave up trying to display multiple dimenions simultaneously and created simple bar charts instead.

%%  was use of pen and paper to collate the values, which was often faster and resulted in less observed frustration. While the majority of users did not attempt anything complex with Excel (in fact only one user used charts), there was some advanced use of Excel observed.
